---
title: "p8105_hw2_wq2160"
author: "Wenshan Qu (wq2160)"
date: "10/8/2021"
output: github_document
---

```{r include = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Read and clean the Mr. Trash Wheel data set.

```{r}
trash_wheel = 
  read_excel("./data_1/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 1, range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  drop_na(month) %>% 
  mutate(
    sports_balls = as.integer(round(sports_balls))
  )

trash_wheel
```

Read and clean the precipitation data for 2018.

```{r}
pre_18 = 
  read_excel("./data_1/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
    year = c("2018")
  )

pre_18
```

Read and clean the precipitation data for 2019.

```{r}
pre_19 = 
  read_excel("./data_1/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
    year = c("2019")
  )

pre_19
```

Combine precipitation data sets of 2018 and 2019.

```{r}
pre_1819 = 
  left_join(pre_18, pre_19, by = "month") %>% 
  mutate(
    month = month.name[month]
  ) %>% 
  relocate(month, year.x, total.x, year.y, total.y)

pre_1819
```

The **number of observation** in trash wheel dataset is `nrow(trash_wheel)`; in 2018 Precipitation is `nrow(pre_18)`, and in 2019 Precipitation is `nrow(pre_19)`, in combination set of 2018 and 2019 precipitation is `nrow(pre_1819)`.

The **key variables** in trash wheel is `names(trash_wheel)`, in 2018 Precipitation is `names(pre_18)`, in 2019 Precipitation is `names(pre_19)`, and in combination set of 2018 and 2019 precipitation is `names(pre_1819)`.

The **total precipitation** in 2018 is `sum(pull(pre_18, total))`, in 2019 is `sum(pull(pre_19, total))`. And the information get from `pre_1819` is that the precipitation amount in 2019 is much lower than 2018, and this trend is extremely obvious during summer time (from April to September).

The **median number of sports balls** in a dumpster in 2019 is `filter(trash_wheel, year == "2019") %>% pull(sports_balls) %>% median()`.


## Problem 2

Clean and modify the pols-month data set.

```{r}
pols_df = 
  read.csv("./data_2/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day)
  ) %>% 
  mutate(
    month = month.name[month] ## replace month number with month name 
  ) %>% 
  select(- day) %>% 
  relocate(year, month, prez_gop, prez_dem) %>% 
  mutate(
    prez_gop = na_if(prez_gop, 0),
    prez_dem = na_if(prez_dem, 0)  ## convert value "0" to NA for future manipulation
  ) %>% 
  pivot_longer(
    prez_gop:prez_dem,
    names_to = "name", 
    names_prefix = "prez_",
    values_to = "president"
  ) %>% 
  drop_na() %>% ## only remain one row (whether dem or gop) for each month of year
  select(- president) %>% ## clean the 0 and 1 value because it's meaningless, only reserve the "dem" or "gop" value
  rename(president = name) %>% ## rename the column containing what we need as president
  select(year, month, president, everything())

pols_df
```

Clean and modify snp data set.

```{r message = FALSE}
library(lubridate) ## we use lubridate package to extract the date information
```

```{r}
snp_df = 
  read.csv("./data_2/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    month = month(mdy(date)),
    year = year(mdy(date)), ## use lubridate functions to extract month and year
    year = ifelse(year >= 2050, year - 100, year) ## lubridate will recognize two-digit old years (i.e. 50 ~ 68) as 2050 ~2068 instead of 1950 ~ 1968, so we manually correct it.
  ) %>% 
  select(year, month, close) %>% 
  mutate(
    month = month.name[month]
  ) %>% 
  arrange(year, month)

snp_df
```

Clean and modify unemployment data.

```{r}
unemployment_df = 
  read.csv("./data_2/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
  mutate(
    month = recode(month, "jan" = 1, "feb" = 2, "mar" = 3, "apr" = 4, "may" = 5, "jun" = 6, "jul" = 7, "aug" = 8, "sep" = 9, "oct" = 10, "nov" = 11, "dec" = 12),
    month = month.name[month]
  )

unemployment_df
```

Join the datasets.

```{r}
pols_snp = 
  left_join(pols_df, snp_df, by = c("year", "month"))
pols_snp_unemploy = 
  left_join(pols_snp, unemployment_df, by = c("year", "month")) %>% 
  arrange(year, month)

pols_snp_unemploy
head(pols_snp_unemploy)
tail(pols_snp_unemploy)
```


**pols-month** data set contains the number of national politicians who are democratic or republican at any given time, such as at a given month and year, whether the president belongs to democratic or republican party, and the number of republican or democratic governors, senators and representatives.

**snp** data set presents Standard & Poor’s stock market index, which shows the closing values of the S&P stock index on the associated date.

**unemployment** data set contains the percentage of unemployment at a given month and year.

**The resulting data set** contains 822 rows and 11 columns, and the year range is from 1947 to 2015, and the names of key variables could be seen by `names(pols_snp_unemploy)`, which contains "year", "month", "president", "gov_gop", "sen_gop", "rep_gop", "gov_dem", "sen_dem", "rep_dem", "close" and "unemployment". Note that the `unemployment` data set range from 1948 to 2015, and the `snp` range from 1950 to 2015.


## Problem 3

Load and tidy the data.

```{r}
bbname_df = 
  read.csv("./data_3/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  rename(childs_first_name = child_s_first_name) %>% 
  mutate(
    ethnicity = str_to_lower(ethnicity),
    childs_first_name = str_to_title(childs_first_name), ## change names to "Abcde" format
    year_of_birth = as.integer(year_of_birth),
    gender = str_to_lower(gender),
    gender = as.factor(gender),
    ethnicity = recode(ethnicity,
                       "asian and pacific islander" = "asian and paci",
                       "black non hispanic" = "black non hisp",
                       "white non hispanic" = "white non hisp")
  ) %>% 
  distinct()
```

